{\bf Practical Concurrency Testing}

### 2018-10-17 thesis/defense/snakefight.tex
last maybe

	Testing approaches
	\begin{itemize}
		\item Unit tests: only 1 ``most likely'' interleaving
		\item Stress tests: exposes ``less likely'' interleavings (if lucky), % to expose "less likely" ones
			cannot verify
		\item Verification: steep learning curve, prohibitive runtime
			%still no widespread adoption % -- why?
	\end{itemize}
\end{frame}
\begin{frame}{Practical Concurrency Testing}
	%Goals
	\textbf{Landslide}, this thesis's solution
	\begin{itemize}
		\item No randomness, despite nondeterminism
		\item Confirm when you fixed bugs, not just find them % not 'verify'
			% TODO: make alnguage match, "Automatic so and so"
		\item User-friendly enough for students
		% \item Test wide variety of types of concurrency % txn
			% TODO : figure out how to ref 
		% ????
		\item Combines with human intuition to verify tests neither could do alone
	\end{itemize}
	%
	% enter landslide
	% - test yr nondeterministic program without randomness/unpredictability
	% - provide formal verification guarantees so you know when you actually fixed that bug
	%   - (well, subject to assumptions about the test case you're running under, etc etc)
	% - user friendly enough even for ze studence
	% 
	% say: "What if there was a program that... ...I'm here to tell you that I've built such a program,
	% it's called Landslide, and why you should let me graduate for it" (figure out smth more diplomatic here...)
\end{frame}

@@

	Too much manual intervention %for non-expert use
	\begin{itemize}
			% TODO: remember to say, "that lets you choose more intelligently, but still by hand, yr PPs
		\item State space size estimation {\em [Simsa '12]}
			% TODO: i just switched the order of these; see how that feels
		\item Manual kernel annotations {\em [Blum '12]} %{\tiny that's me!}
			\pause
		\item Widespread adoption requires {\bf fully automated instrumentation}...
	\end{itemize}
	\pause
	\linegap

@@

\begin{frame}{Thesis Statement}
	% data race pps (as well as txn equivalence)
	Combining %both
	theoretically-founded automatic reduction techniques
	% iterative deepening
	% (say smth like: "landslide also has a wealth of other user-tunable heuristics
	% for expert users, which you can read about in the document; for this talk i will focus just on the CPU budget thing")
	% XXX: need to prepare for the question -- "what evidence do you have that users actually want this cpu budget stuff",
	% ",or, that it helps vs just running something unlimited vs a normal icb test"
	% (well, for the 2nd half of the Q, i can talk about how ID outperforms ICB on verficiations)
	and user-informed heuristic ones,
	stateless model checking can sufficiently mitigate exponential explosion to be
	% education
	practical for inexperienced users
	%a practical testing technique for inexperienced users
	% htm
	% TODO: this last part you just added for this talk; does the thesis version of the statmeent need to include too?
	and real-world programs alike.
\end{frame}

\begin{frame}{Thesis Statement}
	% data race pps (as well as txn equivalence)
	\hilight{gray}{Combining %both
	theoretically-founded automatic reduction techniques
	% iterative deepening
	% (say smth like: "landslide also has a wealth of other user-tunable heuristics
	% for expert users, which you can read about in the document; for this talk i will focus just on the CPU budget thing")
	% XXX: need to prepare for the question -- "what evidence do you have that users actually want this cpu budget stuff",
	% ",or, that it helps vs just running something unlimited vs a normal icb test"
	% (well, for the 2nd half of the Q, i can talk about how ID outperforms ICB on verficiations)
	and user-informed heuristic ones, stateless model checking...}
	\begin{itemize}
		\item \hilight{sect-quicksand}{...can sufficiently mitigate exponential explosion...}\xspace
		% education
		%\item ...to be a practical testing technique for inexperienced users...
		\item \hilight{sect-410}{...to be practical for inexperienced users...}\xspace
		% htm
		% TODO: this last part you just added for this talk; does the thesis version of the statmeent need to include too?
		\item \hilight{sect-htm}{...and real-world programs alike.}\xspace
	\end{itemize}
\end{frame}

@@

% TODO put a pict of mario-man, i.e maximal state space

\begin{frame}{Total Verification}
	\textbf{Convergence theorem:} Testing maximal state space, after detecting all data races, $\equiv$ preempting everywhere!
	\begin{itemize}
		\item $\forall$ arbitrary buggy interleaving, $\exists$ equivalent interleaving % say 'under DPOR' i guess?
			using only sync/data-race PPs, \& Quicksand will eventually reach it.
			%\begin{itemize}
			%	\item ({\em equivalence} defined under DPOR {\em [Flanagan '05]})
			%\end{itemize}
	\end{itemize}
	\linegap

@@

	Each tested for 10 CPU-hours with
	\begin{itemize}
		\item Quicksand % parallelized 10-fold
			%, finding data-race candidates with...
			%\begin{itemize}
			%	\item ...``Pure'' Happens-Before analysis {\em [Lamport '78]}
			%	\item ...``Limited'' Happens-Before analysis {\em [O'Callahan '03]}
			%\end{itemize}
		\item Stand-alone Landslide using ICB {\em [Musuvathi '08]}, preempting at...
			\begin{itemize}
				\item ...synchronization APIs only {\em [Simsa '10]}
				\item ...all shared memory accesses
					%{\em [Holzmann '97]} % not spin
					{\em [Yang '08]} % inspect
			\end{itemize}
	\end{itemize}
\end{frame}

@@

	Ad-hoc synchronization (e.g. {\tt while (!ready) yield();})
	\begin{itemize}
		\item Not an infinite loop bug!
		\item Heuristically ``block'' threads when 10+ {\tt yield()}s %({\tt xchg}, etc.)
		\item Existing memory analysis %(DPOR)
			on loop condition to unblock
			\begin{itemize}
					% TODO: cut this and just say it out loud?
				\item (e.g., what other thread writes to {\tt ready}?)
				% it's kinda... idk how to critique its unsoundness.
				% they self-admit to the unsoundness but like, leaving the cycle in,
				% is the wrong approach to begin with
					% TODO: move this to bonus slides, save a min of expl
				%\item Prior work {\em [Coons '13]} likely to leave threads blocked too long
			\end{itemize}
			\pause
		\item ``Ok, what if I just {\tt yield()} 10 times because I feel like it?''
			\begin{itemize}
				\item If deadlock suspected, try waking up to avoid false positives
			\end{itemize}
		%\item Check for false-positive blocking to avoid false-positive deadlocks
	\end{itemize}

	% Unlike kernels [Blum '12], P2 can be......
	% TODO talk abt yield loops, 
	% compare to prior technique - delay abounding - which sucks bc (verify this by reading BPOR) it doesn't actually
	% wake up the other thread at the right time (using memory conflix)

@@

% do even need this slide?
%\begin{frame}{Evaluation Questions}
%	asfda % TODO
%	% submit better projects?
%	% get better grades?
%	% like the experience?
%\end{frame}

@@

	Pintos (U. Chicago)
	\begin{itemize}
		\item TA %(Kevin Zhao)
			ran Landslide on {\em threads} and {\em userprog} submissions
		\item Returned preemption traces to students after deadline
		\item Surveyed students on ease of debugging, etc.
	\end{itemize}
	% TODO: irb stuff in a bonus slide
\end{frame}

@@

	Bug-finding
	\begin{itemize}
		\item CMU: 115 deterministic, 122 concurrency bugs in 75/103 P2s
			%\begin{itemize}
			%	\item 3 false positives (self-reported)
			%\end{itemize}
		\item U. Chicago: 3 deterministic, 4 concurrency bugs in 3/4 kernels
			% TODO: what to do abt this bullet pts
			%\pause
		\item PSU: (no log collection) %due to IRB issues)
		% \item Case study tej
	\end{itemize}
\end{frame}

@@

%\begin{frame}{Survey Results} % (happiness value)}
%	\begin{center}
%		\includegraphics[width=\textwidth]{../survey5.pdf} % easy to understand?
%	\end{center}
%\end{frame}

@@

\begin{frame}{Summary}
	Landslide helps students find \& fix bugs during projects
	% TODO: should this be moved ot the earlier bug finding slide
	\begin{itemize}
		\item 73\% of CMU users found $\ge 1$ bug
		\item 56\% of CMU users fixed $\ge 1$ bug
		\item 34\% of CMU users fixed all found bugs
	\end{itemize}
	\linegap

@@

	stateless model checking
	\hilight{sect-quicksand}{can sufficiently mitigate exponential explosion}\xspace
	\hilight{sect-410}{to be practical for inexperienced users}\xspace
	\hilight{sect-htm}{and real-world programs alike.}\xspace
\end{frame}
\begin{frame}{Conclusion}
	%\hilight{gray}{Combining %both
	%{\bf theoretically-founded automatic reduction techniques}}
	Combining %both
	{\bf theoretically-founded automatic reduction techniques}
	\begin{itemize}
		\item \hilight{sect-quicksand}{Verification with data-race preemption points}
		\item \hilight{sect-410}{Student yield-loop blocking} % student (adj). from "studence"
		\item \hilight{sect-htm}{Equivalent emulation of HTM semantics}
	\end{itemize}
	% TODO: incremental reveal
	%\hilight{gray}{and {\bf user-informed heuristic ones},}
	and {\bf user-informed heuristic ones},
	\begin{itemize}
		\item \hilight{sect-quicksand}{Quicksand's CPU budget argument}
		\item \hilight{sect-410}{Exposing data-race reports to students}
		%\item Encouraging students to exercise their intuition % ??
		\item \hilight{sect-htm}{Abstraction reduction for HTM primitives}
	\end{itemize}
	\linegap

@@

\subsection{Prior Work}
\breakslide{\Large \bf Bonus Slides - Prior Work}

@@

@@

\begin{frame}{Statistical Significance}
	Compared %users' vs. non-users'
	grade distributions with k-sample Anderson-Darling {\em [A., D. '52]}
	\linegap

	Control 1: non-users from {\bf same semester}
	\begin{itemize}
		\item {\bf P2} grade distributions differ with $p = 0.036 < 0.05$
		\item {\bf P3} grade distributions differ with $p = 0.034 < 0.05$
			\pause
		\item Possible bias: stronger students volunteer for Landslide in first place
	\end{itemize}
	\pause
	\linegap

	Control 2: non-users from {\bf past semesters} (when Landslide not offered)
	\begin{itemize}
		\item {\bf P2} grade distributions differ with $p = 0.671$...
		\item {\bf P3} grade distributions differ with $p = 0.145$...
			\pause
		\item Possible bias: grading criteria may drift across semesters
			% FIXME does this bullet belong in this sub list
		\item Possible bias: maybe Landslide finds bugs the official tests don't
	\end{itemize}
\end{frame}

@@

%\begin{frame}{HTM: Sources of Benchmarks}
%	\begin{itemize}
%		%\item Shamay, https://github.com/cmpxchg16/tsx.me, 2014
%		\item Dehesa-Azuara and Stanley, http://www.contrib.andrew.cmu.edu/\textasciitilde{}mdehesaa/, 2016
%		%\item Chapman et al., Hybrid STM/HTM for nested transactions on OpenJDK, OOPSLA 2016
%		%\item Wang et al., Eunomia: Scaling Concurrent Search Trees under Contention Using HTM, PPoPP 2017
%			% TODO: add, that one lock thingy
%		\item vamsikc, TSX LevelDB implementation, https://github.com/vamsikc/leveldb-tsx/, 2014
%	\end{itemize}
%\end{frame}

@@

\begin{frame}{Retry Reduction}
	% TODO: split this into a step by step
	\begin{center}
		\includegraphics[width=\textwidth]{../retry-sets.pdf}
	\end{center}
\end{frame}

@@

			% TODO: check these again, will 2,4 even complete on mutex
			{\tt lock}({\tt spinlock})      & 1     & 2     & 4     & 28    \\ % 2,2
			{\tt lock}({\tt spin\_fixed})   & 1     & 2     & 4     & 28    \\ % 2,2 or 4,1
			{\tt lock}({\tt mutex})	 & 1     & 2     & 2     & 30    \\ % 2,3 gives 30; 2,4 would give 40
			% not putting lock_fast because it like, doesn't have a "state space" per se
		\end{tabular}
	\end{center}
	\linegap

@@

%% including this stuff in the talk now so this bonus slide is frivolous
%% TODO: slide number c.c
%\begin{frame}{Don't you need \texttt{if (mutex\_is\_locked(\&count\_lock)) \_xabort();} in that transaction on slide 42?}
%	\begin{center}
%	Yes.
%	\end{center}
%\end{frame}

@@

\begin{frame}{References}
	\footnotesize
	\begin{itemize}
		\item {\bf [A., D. '52]}: T. Anderson and D. Darling.
			Asymptotic theory of certain ``goodness-of-fit'' criteria based on stochastic processes.
			Annals of Statistics, 1952.
		\item {\bf [Lamport '78]}:
			Leslie Lamport. Time, clocks, and the ordering of events in a distributed system.
			Communications of the ACM, 1978.
		\item {\bf [Godefroid '97]}: Patrice Godefroid.
			VeriSoft. A tool for the automatic analysis of concurrent reactive software. CAV 1997.
		\item {\bf [Savage '97]}: Stefan Savage.
			Eraser: a dynamic data race detector for multi-threaded programs. SOSP 1997.
		%\item {\bf [Holzmann '97]}: The model checker SPIN. TOSE 1997.
		\item {\bf [Engler '03]}: Dawson Engler and Ken Ashcraft.
			RacerX: effective, static detection of race conditions and deadlocks. SOSP 2003.
		\item {\bf [O'Callahan '03]}: Robert O'Callahan and Jong-Deok Choi.
			Hybrid dynamic data race detection. PPoPP '03.
		\item {\bf [Flanagan '05]}: Cormac Flanagan and Patrice Godefroid. Dynamic partial-order reduction for
			model checking software. POPL 2005.
		\item {\bf [Musuvathi '08]}: Madanlal Musuvathi et al. Finding and Reproducing Heisenbugs in Concurrent
			Programs. OSDI 2008.
		\item {\bf [Yang '08]}: Yu Yang et al. Efficient stateful DPOR. SPIN 2008.
		\item {\bf [Klein '09]}: Gerwin Klein et al. seL4: formal verif. of an OS kernel. SOSP 2009.
	\end{itemize}
\end{frame}
\begin{frame}{References}
	\footnotesize
	\begin{itemize}
		%\item {\bf [Burckhardt '10]}: A randomized scheduler with probabilistic guarantees of finding bugs. ASPLOS 2010.
		\item {\bf [Simsa '12]}: Jiri Simsa. Runtime Estimation and Resource Allocation for
			Concurrency Testing. Tech report (CMU-PDL-12-113). December 2012.
		\item {\bf [Blum '12]}: Ben Blum. Landslide: Systematic Dynamic Race Detection in Kernel Space.
			MS thesis (CMU-CS-12-118). May 2012.
		\item {\bf [Kasicki '12]}: Baris Kasicki. Data races vs. data race bugs: telling the difference with Portend. ASPLOS '12.
		\item {\bf [Simsa '13]}: Systematic and Scalable Testing of Concurrent Programs.
			Ph.D. thesis (CMU-CS-13-133). December 2013.
		%\item {\bf [Coons '13]}: Katherine Coons et al. Bounded POR. OOPSLA 2013.
		\item {\bf [Abdulla '14]}: Parosh Abdulla et al. Optimal DPOR. POPL 2014.
		\item {\bf [Matsakis '14]}: Nicholas Matsakis et al. The Rust Language. HILT 2014.
		\item {\bf [vamsikc '14]}: leveldb-tsx. https://github.com/vamsikc/leveldb-tsx/, 2014.
		%\item {\bf [Huang '15]}: Jeff Huang. Stateless MCing concurrent programs with maximal causality reduction. PLDI '15.
		\item {\bf [Jensen '15]}: Stateless MCing of Event-Driven Applications. OOPSLA 2015.
		\item {\bf [Dehesa-Azuara '16]}: Hardware transactional memory with intel's TSX.
			http://www.contrib.andrew.cmu.edu/\textasciitilde{}mdehesaa/, 2016.
	\end{itemize}
\end{frame}

>>>>

	Testing approaches
	\begin{itemize}
		\item Unit tests: only 1 ``most likely'' interleaving
		\item Stress tests: exposes ``less likely'' interleavings (if lucky), % to expose "less likely" ones
			cannot verify
		\item Verification: steep learning curve, prohibitive runtime
			%still no widespread adoption % -- why?
	\end{itemize}
\end{frame}
\begin{frame}{Practical Concurrency Testing}
	%Goals
	\textbf{Landslide}, this thesis's solution
	\begin{itemize}
		\item No randomness, despite nondeterminism
		\item Confirm when you fixed bugs, not just find them % not 'verify'
		\item User-friendly enough for students
		% \item Test wide variety of types of concurrency % txn
		% ????
		\item Combines with human intuition to verify tests neither could do alone
	\end{itemize}
	%
	% enter landslide
	% - test yr nondeterministic program without randomness/unpredictability
	% - provide formal verification guarantees so you know when you actually fixed that bug
	%   - (well, subject to assumptions about the test case you're running under, etc etc)
	% - user friendly enough even for ze studence
	% 
	% say: "What if there was a program that... ...I'm here to tell you that I've built such a program,
	% it's called Landslide, and why you should let me graduate for it" (figure out smth more diplomatic here...)
\end{frame}

@@

	Too much manual intervention %for non-expert use
	\begin{itemize}
			% TODO: remember to say, "that lets you choose more intelligently, but still by hand, yr PPs
		\item State space size estimation {\em [Simsa '12]}
		\item Manual kernel annotations {\em [Blum '12]} %{\tiny that's me!}
			\pause
		\item Widespread adoption requires {\bf fully automated instrumentation}...
	\end{itemize}
	\pause
	\linegap

@@

\begin{frame}{Thesis Statement}
	% data race pps (as well as txn equivalence)
	Combining %both
	theoretically-founded automatic reduction techniques
	% iterative deepening
	% (say smth like: "landslide also has a wealth of other user-tunable heuristics
	% for expert users, which you can read about in the document; for this talk i will focus just on the CPU budget thing")
	% XXX: need to prepare for the question -- "what evidence do you have that users actually want this cpu budget stuff",
	% ",or, that it helps vs just running something unlimited vs a normal icb test"
	% (well, for the 2nd half of the Q, i can talk about how ID outperforms ICB on verficiations)
	and user-informed heuristic ones,
	stateless model checking can sufficiently mitigate exponential explosion to be
	% education
	practical for inexperienced users
	%a practical testing technique for inexperienced users
	% htm
	and real-world programs alike.
\end{frame}

\begin{frame}{Thesis Statement}
	% data race pps (as well as txn equivalence)
	\hilight{gray}{Combining %both
	theoretically-founded automatic reduction techniques
	% iterative deepening
	% (say smth like: "landslide also has a wealth of other user-tunable heuristics
	% for expert users, which you can read about in the document; for this talk i will focus just on the CPU budget thing")
	% XXX: need to prepare for the question -- "what evidence do you have that users actually want this cpu budget stuff",
	% ",or, that it helps vs just running something unlimited vs a normal icb test"
	% (well, for the 2nd half of the Q, i can talk about how ID outperforms ICB on verficiations)
	and user-informed heuristic ones, stateless model checking...}
	\begin{itemize}
		\item \hilight{sect-quicksand}{...can sufficiently mitigate exponential explosion...}\xspace
		% education
		%\item ...to be a practical testing technique for inexperienced users...
		\item \hilight{sect-410}{...to be practical for inexperienced users...}\xspace
		% htm
		\item \hilight{sect-htm}{...and real-world programs alike.}\xspace
	\end{itemize}
\end{frame}

@@

\begin{frame}{Total Verification}
	Soundness theorem: Testing maximal state space, after detecting all data races, $\equiv$ preempting everywhere!
	\begin{itemize}
		\item $\forall$ arbitrary buggy interleaving, $\exists$ equivalent interleaving % say 'under DPOR' i guess?
			using only sync/data-race preemption points %, \& Quicksand will eventually reach it.
			%\begin{itemize}
			%	\item ({\em equivalence} defined under DPOR {\em [Flanagan '05]})
			%\end{itemize}
	\end{itemize}
	\linegap

@@

	Each tested for 10 CPU-hours with
	\begin{itemize}
		\item Quicksand % parallelized 10-fold
			%, finding data-race candidates with...
			%\begin{itemize}
			%	\item ...``Pure'' Happens-Before analysis {\em [Lamport '78]}
			%	\item ...``Limited'' Happens-Before analysis {\em [O'Callahan '03]}
			%\end{itemize}
		\item Stand-alone Landslide %using ICB {\em [Musuvathi '08]},
			preempting at...
			\begin{itemize}
				\item ...synchronization APIs only {\em [Simsa '13]}
				\item ...all shared memory accesses
					%{\em [Holzmann '97]} % not spin
					{\em [Yang '08]} % inspect
			\end{itemize}
	\end{itemize}
\end{frame}

@@

	Ad-hoc synchronization (e.g. {\tt while (!ready) yield();})
	\begin{itemize}
		\item Not an infinite loop bug!
		\item Heuristically ``block'' threads when 10+ {\tt yield()}s %({\tt xchg}, etc.)
		\item Existing memory analysis %(DPOR)
			on loop condition to unblock
			\begin{itemize}
				\item (e.g., what other thread writes to {\tt ready}?)
				% it's kinda... idk how to critique its unsoundness.
				% they self-admit to the unsoundness but like, leaving the cycle in,
				% is the wrong approach to begin with
				%\item Prior work {\em [Coons '13]} likely to leave threads blocked too long
			\end{itemize}
			\pause
		\item ``Ok, what if I just {\tt yield()} 10 times because I feel like it?''
			\begin{itemize}
				\item If deadlock suspected, try waking up to avoid false positives
			\end{itemize}
		%\item Check for false-positive blocking to avoid false-positive deadlocks
	\end{itemize}

@@

@@

	Pintos (U. Chicago)
	\begin{itemize}
		\item TA %(Kevin Zhao)
			ran Landslide on {\em threads} and {\em userprog} submissions
		\item Returned preemption traces to students after deadline
		\item Surveyed students on ease of debugging, etc.
	\end{itemize}
\end{frame}

@@

	Bug-finding
	\begin{itemize}
		\item CMU: 115 deterministic, 122 concurrency bugs in 75/103 P2s
		\item U. Chicago: 3 deterministic, 4 concurrency bugs in 3/4 kernels
		\item PSU: (no log collection) %due to IRB issues)
		% \item Case study tej
	\end{itemize}
\end{frame}

@@

@@

\begin{frame}{Summary}
	Landslide helps students find \& fix bugs during projects
	\begin{itemize}
		\item 73\% of CMU users found $\ge 1$ bug
		\item 56\% of CMU users fixed $\ge 1$ bug
		\item 34\% of CMU users fixed all found bugs
	\end{itemize}
	\linegap

@@

	%stateless model checking
	%\hilight{sect-quicksand}{can sufficiently mitigate exponential explosion}\xspace
	%\hilight{sect-410}{to be practical for inexperienced users}\xspace
	%\hilight{sect-htm}{and real-world programs alike.}\xspace
	\hilight{gray}{stateless model checking}\xspace
	\hilight{sect-pastel-quicksand}{can sufficiently mitigate exponential explosion}\xspace
	\hilight{sect-pastel-410}{to be practical for inexperienced users}\xspace
	\hilight{sect-pastel-htm}{and real-world programs alike.}\xspace
\end{frame}
\begin{frame}{Conclusion}
	%\hilight{gray}{Combining %both
	%{\bf theoretically-founded automatic reduction techniques}}
	Combining %both
	{\bf theoretically-founded automatic reduction techniques}
	\begin{itemize}
		\item \hilight{sect-quicksand}{Verification with data-race preemption points}
		\item \hilight{sect-410}{Student yield-loop blocking} % student (adj). from "studence"
		\item \hilight{sect-htm}{Equivalent emulation of HTM semantics}
	\end{itemize}
	%\hilight{gray}{and {\bf user-informed heuristic ones},}
	and {\bf user-informed heuristic ones},
	\vspace{4.82em}
	\linegap

	\hilight{gray}{stateless model checking}\xspace
	\hilight{sect-pastel-quicksand}{can sufficiently mitigate exponential explosion}\xspace
	\hilight{sect-pastel-410}{to be practical for inexperienced users}\xspace
	\hilight{sect-pastel-htm}{and real-world programs alike.}\xspace
\end{frame}
\begin{frame}{Conclusion}
	%\hilight{gray}{Combining %both
	%{\bf theoretically-founded automatic reduction techniques}}
	Combining %both
	{\bf theoretically-founded automatic reduction techniques}
	\begin{itemize}
		\item \hilight{sect-quicksand}{Verification with data-race preemption points}
		\item \hilight{sect-410}{Student yield-loop blocking} % student (adj). from "studence"
		\item \hilight{sect-htm}{Equivalent emulation of HTM semantics}
	\end{itemize}
	%\hilight{gray}{and {\bf user-informed heuristic ones},}
	and {\bf user-informed heuristic ones},
	\begin{itemize}
		\item \hilight{sect-quicksand}{Quicksand's CPU budget argument}
		\item \hilight{sect-410}{Exposing data-race reports to students}
		%\item Encouraging students to exercise their intuition % ??
		\item \hilight{sect-htm}{Abstraction reduction for HTM primitives}
	\end{itemize}
	\linegap

@@

\subsection{Background}
\breakslide{\Large \bf Bonus Slides - Background}

@@

\begin{frame}{Verification Soundness}
	\textbf{Reachability} of preemption points
	\begin{itemize}
		\item {\tt mutex\_lock()}, {\tt mutex\_unlock()}, {\tt yield()}, etc, are reachable.
		\item A data race PP is reachable if it will be identified while preempting on other reachable PPs.
	\end{itemize}
	\linegap

	{\bf Convergence}: Quicksand will eventually detect all reachable data races.
	\linegap

	{\bf Verification}: $\forall$ buggy interleavings, $\exists$ an equivalent interleaving exposing the same bug
	that switches threads only on reachable PPs.
	\linegap

	No bugs found testing reachable PPs $\rightarrow$ no buggy interleaving can exist
\end{frame}

@@

\begin{frame}{Grades - Statistical Significance}
	Compared %users' vs. non-users'
	grade distributions with k-sample Anderson-Darling {\em [A., D. '52]}
	\linegap

	Control 1: non-users from {\bf same semester}
	\begin{itemize}
		\item {\bf P2} grade distributions differ with $p = 0.036 < 0.05$
		\item {\bf P3} grade distributions differ with $p = 0.034 < 0.05$
			%\pause
		\item Possible bias: stronger students volunteer for Landslide in first place
	\end{itemize}
	%\pause
	\linegap

	Control 2: non-users from {\bf past semesters} (when Landslide not offered)
	\begin{itemize}
		\item {\bf P2} grade distributions differ with $p = 0.671$...
		\item {\bf P3} grade distributions differ with $p = 0.145$...
			%\pause
		\item Possible bias: grading criteria may drift across semesters
			% FIXME does this bullet belong in this sub list
		\item Possible bias: maybe Landslide finds bugs the official tests don't
	\end{itemize}
\end{frame}

\begin{frame}{Survey Results} \begin{center} \includegraphics[width=\textwidth]{../survey1.pdf} \end{center} \end{frame}
\begin{frame}{Survey Results} \begin{center} \includegraphics[width=\textwidth]{../survey2.pdf} \end{center} \end{frame}
\begin{frame}{Survey Results} \begin{center} \includegraphics[width=\textwidth]{../survey3.pdf} \end{center} \end{frame}
\begin{frame}{Survey Results} \begin{center} \includegraphics[width=\textwidth]{../survey4.pdf} \end{center} \end{frame}
\begin{frame}{Survey Results} \begin{center} \includegraphics[width=\textwidth]{../survey5.pdf} \end{center} \end{frame}
\begin{frame}{Survey Results} \begin{center} \includegraphics[width=\textwidth]{../survey7.pdf} \end{center} \end{frame}

@@

%\begin{frame}{HTM: Sources of Benchmarks}
%	\begin{itemize}
%		%\item Shamay, https://github.com/cmpxchg16/tsx.me, 2014
%		\item Dehesa-Azuara and Stanley, http://www.contrib.andrew.cmu.edu/\textasciitilde{}mdehesaa/, 2016
%		%\item Chapman et al., Hybrid STM/HTM for nested transactions on OpenJDK, OOPSLA 2016
%		%\item Wang et al., Eunomia: Scaling Concurrent Search Trees under Contention Using HTM, PPoPP 2017
%		\item vamsikc, TSX LevelDB implementation, https://github.com/vamsikc/leveldb-tsx/, 2014
%	\end{itemize}
%\end{frame}

@@

\begin{frame}{Retry Set Reduction}
	\begin{center}
		\includegraphics[width=\textwidth]{../retry-sets.pdf}
	\end{center}
\end{frame}

@@

			{\tt lock}({\tt spinlock})      & 1     & 2     & 4     & 28    \\ % 2,2
			{\tt lock}({\tt spin\_fixed})   & 1     & 2     & 4     & 28    \\ % 2,2 or 4,1
			{\tt lock}({\tt mutex})	 & 1     & 2     & 2     & 30    \\ % 2,3 gives 30; 2,4 would give 40
			% not putting lock_fast because it like, doesn't have a "state space" per se
		\end{tabular}
	\end{center}
	\linegap

@@

\begin{frame}{HTM \& Weak Memory Reorderings}
	TSO (x86): Stores may be reordered after subsequent loads
	\linegap

	\begin{center}
	\begin{tabular}{l}
		\texttt{if (\hilight{darkorange}{\_xbegin}() == \_XBEGIN\_STARTED) \{} \\
		\texttt{~~~~if (prevent\_txn)} \\
		\texttt{~~~~~~~~\_xabort();} \\
		\texttt{~~~~count++;} \\
		\texttt{~~~~\hilight{darkblue}{\_xend}();} \\
		\texttt{\} else \{} \\
		\texttt{~~~~\hilight{darkorange}{mutex\_lock}(\&m);} \\
		\texttt{~~~~prevent\_txn = 1;} \\
		\texttt{~~~~\hilight{red}{// XXX barrier needed here}} \\
		%\texttt{~~~~// \_\_sync\_synchronize(); // uncomment to solve the bug} \\
		\texttt{~~~~count++;} \\
		\texttt{~~~~prevent\_txn = 0;} \\
		\texttt{~~~~\hilight{darkblue}{mutex\_unlock}(\&m);} \\
		\texttt{\}} \\
	\end{tabular}
	\end{center}
\end{frame}

@@

\begin{frame}{References}
	\footnotesize
	\begin{itemize}
		\item {\bf [A., D. '52]}: T. Anderson and D. Darling.
			Asymptotic theory of certain ``goodness-of-fit'' criteria based on stochastic processes.
			Annals of Statistics, 1952.
		\item {\bf [Lamport '78]}:
			Leslie Lamport. Time, clocks, and the ordering of events in a distributed system.
			Communications of the ACM, 1978.
		\item {\bf [Godefroid '97]}: Patrice Godefroid.
			VeriSoft. A tool for the automatic analysis of concurrent reactive software. CAV 1997.
		\item {\bf [Savage '97]}: Stefan Savage.
			Eraser: a dynamic data race detector for multi-threaded programs. SOSP 1997.
		%\item {\bf [Holzmann '97]}: The model checker SPIN. TOSE 1997.
		%\item {\bf [Engler '03]}: Dawson Engler and Ken Ashcraft.
			RacerX: effective, static detection of race conditions and deadlocks. SOSP 2003.
		\item {\bf [O'Callahan '03]}: Robert O'Callahan and Jong-Deok Choi.
			Hybrid dynamic data race detection. PPoPP '03.
		\item {\bf [Flanagan '05]}: Cormac Flanagan and Patrice Godefroid. Dynamic partial-order reduction for
			model checking software. POPL 2005.
		\item {\bf [Musuvathi '08]}: Madanlal Musuvathi et al. Finding and Reproducing Heisenbugs in Concurrent
			Programs. OSDI 2008.
		\item {\bf [Yang '08]}: Yu Yang et al. Efficient stateful DPOR. SPIN 2008.
	\end{itemize}
\end{frame}
\begin{frame}{References}
	\footnotesize
	\begin{itemize}
		\item {\bf [Klein '09]}: Gerwin Klein et al. seL4: formal verif. of an OS kernel. SOSP 2009.
		%\item {\bf [Burckhardt '10]}: A randomized scheduler with probabilistic guarantees of finding bugs. ASPLOS 2010.
		\item {\bf [Simsa '12]}: Jiri Simsa. Runtime Estimation and Resource Allocation for
			Concurrency Testing. Tech report (CMU-PDL-12-113). December 2012.
		\item {\bf [Blum '12]}: Ben Blum. Landslide: Systematic Dynamic Race Detection in Kernel Space.
			MS thesis (CMU-CS-12-118). May 2012.
		%\item {\bf [Kasicki '12]}: Baris Kasicki. Data races vs. data race bugs: telling the difference with Portend. ASPLOS '12.
		\item {\bf [Simsa '13]}: Systematic and Scalable Testing of Concurrent Programs.
			Ph.D. thesis (CMU-CS-13-133). December 2013.
		%\item {\bf [Coons '13]}: Katherine Coons et al. Bounded POR. OOPSLA 2013.
		\item {\bf [Abdulla '14]}: Parosh Abdulla et al. Optimal DPOR. POPL 2014.
		\item {\bf [Matsakis '14]}: Nicholas Matsakis et al. The Rust Language. HILT 2014.
		\item {\bf [vamsikc '14]}: leveldb-tsx. https://github.com/vamsikc/leveldb-tsx/, 2014.
		%\item {\bf [Huang '15]}: Jeff Huang. Stateless MCing concurrent programs with maximal causality reduction. PLDI '15.
		\item {\bf [Jensen '15]}: Stateless MCing of Event-Driven Applications. OOPSLA 2015.
		\item {\bf [Dehesa-Azuara '16]}: Hardware transactional memory with intel's TSX.
			http://www.contrib.andrew.cmu.edu/\textasciitilde{}mdehesaa/, 2016.
	\end{itemize}
\end{frame}
